### 2021-02-22 architecture.tex
Mention problem P when comparing precedences

Each training example has the form $(P, \PrecBetter, \PrecWorse)$,
where $P = (\Sigma, \mathit{Cl})$ is a problem
and $\PrecBetter, \PrecWorse$ are precedences over $\Sigma$
such that the prover using $\PrecBetter$ solves $P$ in fewer iterations of the saturation loop than with $\PrecWorse$,
denoted as $\Better{\PrecBetter}{\PrecWorse}$.\todo{MS: Could we have $\PrecBetter \prec_P\PrecWorse$?}
\todo{FB: Is this sufficient?}
\todo{FB: Describe the distribution of problems and precedences.}

@@

Let $(P, \PrecBetter, \PrecWorse)$ be a training example ($\Better{\PrecBetter}{\PrecWorse}$).
The precedence cost classifies this example correctly if $C(\PrecBetter) < C(\PrecWorse)$,
or alternatively $C_2(\PrecBetter, \PrecWorse) = C(\PrecBetter) - C(\PrecWorse) < 0$.
We approach this problem as an instance of binary classification with the logistic loss \cite{Mohri2018}:
\todo{MS: First mention of a ``binary classifier''!
We classify the examples, sure, but you did not say it yet.}
\begin{align*}
L(P, \PrecBetter, \PrecWorse)
&= - \log \sigmoid C_2(\PrecBetter, \PrecWorse)
%&= - \log \sigmoid (C(\PrecBetter) - C(\PrecWorse)) \\
%&= - \log \sigmoid \sum_{i=1}^n i (c(\PrecBetter(i)) - c(\PrecWorse(i)))
\end{align*}
\todo{MS: neni tu ted spatne zniminko? Nebo spis uz v definici score? Cekal bych, ze pro positivni examply (coz jsou u nas vsechny)
bude chit $s$ kladne.}

@@

We need to train $c$ in a way that ensures that $C$ is minimized by a good precedence.
On pairs,
$C(\pi)$ should be smaller than $C(\rho)$ iff $\Better{\pi}{\rho}$.

@@

Let $P = (\Sigma, \mathit{Cl})$ be a problem.
Let $\PrecBetter, \PrecWorse$ be two precedences over $\Sigma$.
% that are comparable according to AST.
Let $y = 1$ if $\PrecBetter$ is better than $\PrecWorse$,
denoted as $\Better{\PrecBetter}{\PrecWorse}$,
and $y = 0$ otherwise.

The examples that the recommender will be trained on have the form
$(P, \PrecBetter, \PrecWorse)$, where $P = (\Sigma, \mathit{Cl})$ is a \gls{cnf} problem, and $\PrecBetter, \PrecWorse$ are precedences over $\Sigma$
where $\PrecBetter$ yields a faster proof search than $\PrecWorse$,
\todo{FB: Rephrase?}
denoted as $\Better{\PrecBetter}{\PrecWorse}$.

@@

We train the system on results of \gls{atp} executions on problems with random precedences.
Let $P = (\Sigma, \mathit{Cl})$ be a problem
and let $\PrecBetter, \PrecWorse$ be two precedences over $\Sigma$.
We run an \gls{atp} on $P$ with the two precedences.
If the run with $\PrecBetter$ finds the proof successfully and uses less saturation loop iterations than the run with $\PrecWorse$,
we can be certain that $\PrecBetter$ is a better precedence for $P$ than $\PrecWorse$,
denoted as $\Better{\PrecBetter}{\PrecWorse}$.

@@

The examples that the recommender will be trained on have the form
$(P, \PrecBetter, \PrecWorse)$, where $P = (\Sigma, \mathit{Cl})$ is a \gls{cnf} problem, and $\PrecBetter, \PrecWorse$ are precedences over $\Sigma$
where $\PrecBetter$ yields a faster proof search than $\PrecWorse$,
\todo{FB: Rephrase?}
denoted as $\Better{\PrecBetter}{\PrecWorse}$.

We train a binary classifier that decides,
for a pair of precedences $\PrecBetter, \PrecWorse$,
whether $\Better{\PrecBetter}{\PrecWorse}$.
Since all the training examples are oriented so that $\Better{\PrecBetter}{\PrecWorse}$,
the classification target value for all the samples is $1$.
\todo{FB: Somebody might think that this is a problem because the classifier can simply learn to always say "1". The structure of the classifier ($C(x_1) - C(x_2)$) prevents that. Consider explaining.}
Thus the binary crossentropy loss can be expressed as a simple expression:
$$
L(P, \PrecBetter, \PrecWorse) = - \log \sigmoid (C(\PrecWorse) - C(\PrecBetter))
$$

@@

Let $\PrecBetter, \PrecWorse \in \Perm{n}$ be two symbol precedences over $P$.
Let $\PrecBetter$ yield a faster proof search on $P$ than $\PrecWorse$, denoted as $\Better{\PrecBetter}{\PrecWorse}$.
The precedence cost model can be extended to predict the probability that $\PrecBetter$ is better than $\PrecWorse$:
\todo{Fix this. Ensure the polarity is sound.}
\todo{MS: zatim chybi vysvetleni $\sigmoid$.}
\todo{MS: duvod proc mi v tomhle kontextu moc nesedi mluvit o pravdepodobnosti
je ten, ze bez dalsiho vysvetlovani to vypada, ze si myslime, ze
kdyz pomoci $\sigmoid$ splacnem v podstate libovolne hodnoty do
intervalu $(0,1)$, zecne to automaticky fungovat jako pravdepodobnost.}
\begin{align*}
p(\Better{\PrecBetter}{\PrecWorse} | P)
&= \sigmoid \SquareBracket{C(\PrecWorse | P) - C(\PrecBetter | P)} \\
%&= \sigmoid \SquareBracket{k(n) \sum_{i=1}^n c_i \cdot (\inv{\PrecWorse}_i - \inv{\PrecBetter}_i)} \\
&= \sigmoid \SquareBracket{k(n) \DotProd{c(P)}{\inv{\PrecWorse} - \inv{\PrecBetter}}}
\end{align*}
It is easy to see that $p(\Better{\PrecBetter}{\PrecWorse} | P) > 0.5$ if and only if $C(\PrecWorse | P) > C(\PrecBetter | P)$,
and that the value of $p$ can be increased by increasing $C(\PrecWorse | P)$ or decreasing $C(\PrecBetter | P)$.

We consider the difference $C(\PrecWorse | P) - C(\PrecBetter | P)$
to be the "logit"\todo{MS: Logit je taky tezky pojem, urcite ne automaticky znamy u ATP people.} score of the pair of precedences $\PrecBetter, \PrecWorse$,
and denote it by $C(\Better{\PrecBetter}{\PrecWorse} | P)$.

We define the loss of $c$ on the training sample $(\PrecBetter, \PrecWorse, P)$ as binary cross-entropy
with ground truth $\Better{\PrecBetter}{\PrecWorse}$:

\begin{align*}
L(\PrecBetter, \PrecWorse, P)
&= - \log p(\Better{\PrecBetter}{\PrecWorse} | P) \\
&= - \log \sigmoid \SquareBracket{k(n) \DotProd{c(P)}{\inv{\PrecWorse} - \inv{\PrecBetter}}} \\
&= - \log \sigmoid \SquareBracket{k(n) \sum_{i=1}^n c_i \cdot (\inv{\PrecWorse}_i - \inv{\PrecBetter}_i)}
\end{align*}

The loss is differentiable with respect to the symbol costs:
\begin{align*}
\frac{\partial L}{\partial c_i}
&= -\sigmoid(-C(\Better{\PrecBetter}{\PrecWorse} | P)) \cdot k(n) \cdot (\inv{\PrecWorse}_i - \inv{\PrecBetter}_i) \\
&= (p(\Better{\PrecBetter}{\PrecWorse} | P) - 1) \cdot k(n) \cdot (\inv{\PrecWorse}_i - \inv{\PrecBetter}_i)
\end{align*}

>>>>

Each training example has the form $(P, \PrecBetter, \PrecWorse)$,
where $P = (\Sigma, \mathit{Cl})$ is a problem
and $\PrecBetter, \PrecWorse$ are precedences over $\Sigma$
such that the prover using $\PrecBetter$ solves $P$ in fewer iterations of the saturation loop than with $\PrecWorse$,
denoted as $\Better{\PrecBetter}{\PrecWorse}{P}$.
\todo{FB: Is this sufficient?}
\todo{FB: Describe the distribution of problems and precedences.}

@@

Let $(P, \PrecBetter, \PrecWorse)$ be a training example ($\Better{\PrecBetter}{\PrecWorse}{P}$).
The precedence cost classifies this example correctly if $C(\PrecBetter) < C(\PrecWorse)$,
or alternatively $C_2(\PrecBetter, \PrecWorse) = C(\PrecBetter) - C(\PrecWorse) < 0$.
We approach this problem as an instance of binary classification with the logistic loss \cite{Mohri2018}:
\todo{MS: First mention of a ``binary classifier''!
We classify the examples, sure, but you did not say it yet.}
\begin{align*}
L(P, \PrecBetter, \PrecWorse)
&= - \log \sigmoid C_2(\PrecBetter, \PrecWorse)
%&= - \log \sigmoid (C(\PrecBetter) - C(\PrecWorse)) \\
%&= - \log \sigmoid \sum_{i=1}^n i (c(\PrecBetter(i)) - c(\PrecWorse(i)))
\end{align*}
\todo{MS: neni tu ted spatne zniminko? Nebo spis uz v definici score? Cekal bych, ze pro positivni examply (coz jsou u nas vsechny)
bude chit $s$ kladne.}

@@

We need to train $c$ in a way that ensures that $C$ is minimized by a good precedence.
On pairs,
$C(\pi)$ should be smaller than $C(\rho)$ iff $\Better{\pi}{\rho}{P}$.

@@

Let $P = (\Sigma, \mathit{Cl})$ be a problem.
Let $\PrecBetter, \PrecWorse$ be two precedences over $\Sigma$.
% that are comparable according to AST.
Let $y = 1$ if $\PrecBetter$ is better than $\PrecWorse$,
denoted as $\Better{\PrecBetter}{\PrecWorse}{P}$,
and $y = 0$ otherwise.

The examples that the recommender will be trained on have the form
$(P, \PrecBetter, \PrecWorse)$, where $P = (\Sigma, \mathit{Cl})$ is a \gls{cnf} problem, and $\PrecBetter, \PrecWorse$ are precedences over $\Sigma$
where $\PrecBetter$ yields a faster proof search than $\PrecWorse$,
\todo{FB: Rephrase?}
denoted as $\Better{\PrecBetter}{\PrecWorse}{P}$.

@@

We train the system on results of \gls{atp} executions on problems with random precedences.
Let $P = (\Sigma, \mathit{Cl})$ be a problem
and let $\PrecBetter, \PrecWorse$ be two precedences over $\Sigma$.
We run an \gls{atp} on $P$ with the two precedences.
If the run with $\PrecBetter$ finds the proof successfully and uses less saturation loop iterations than the run with $\PrecWorse$,
we can be certain that $\PrecBetter$ is a better precedence for $P$ than $\PrecWorse$,
denoted as $\Better{\PrecBetter}{\PrecWorse}{P}$.

@@

The examples that the recommender will be trained on have the form
$(P, \PrecBetter, \PrecWorse)$, where $P = (\Sigma, \mathit{Cl})$ is a \gls{cnf} problem, and $\PrecBetter, \PrecWorse$ are precedences over $\Sigma$
where $\PrecBetter$ yields a faster proof search than $\PrecWorse$,
\todo{FB: Rephrase?}
denoted as $\Better{\PrecBetter}{\PrecWorse}{P}$.

We train a binary classifier that decides,
for a pair of precedences $\PrecBetter, \PrecWorse$,
whether $\Better{\PrecBetter}{\PrecWorse}{P}$.
Since all the training examples are oriented so that $\Better{\PrecBetter}{\PrecWorse}{P}$,
the classification target value for all the samples is $1$.
\todo{FB: Somebody might think that this is a problem because the classifier can simply learn to always say "1". The structure of the classifier ($C(x_1) - C(x_2)$) prevents that. Consider explaining.}
Thus the binary crossentropy loss can be expressed as a simple expression:
$$
L(P, \PrecBetter, \PrecWorse) = - \log \sigmoid (C(\PrecWorse) - C(\PrecBetter))
$$

@@

Let $\PrecBetter, \PrecWorse \in \Perm{n}$ be two symbol precedences over $P$.
Let $\PrecBetter$ yield a faster proof search on $P$ than $\PrecWorse$, denoted as $\Better{\PrecBetter}{\PrecWorse}{P}$.
The precedence cost model can be extended to predict the probability that $\PrecBetter$ is better than $\PrecWorse$:
\todo{Fix this. Ensure the polarity is sound.}
\todo{MS: zatim chybi vysvetleni $\sigmoid$.}
\todo{MS: duvod proc mi v tomhle kontextu moc nesedi mluvit o pravdepodobnosti
je ten, ze bez dalsiho vysvetlovani to vypada, ze si myslime, ze
kdyz pomoci $\sigmoid$ splacnem v podstate libovolne hodnoty do
intervalu $(0,1)$, zecne to automaticky fungovat jako pravdepodobnost.}
\begin{align*}
p(\Better{\PrecBetter}{\PrecWorse}{P})
&= \sigmoid \SquareBracket{C(\PrecWorse | P) - C(\PrecBetter | P)} \\
%&= \sigmoid \SquareBracket{k(n) \sum_{i=1}^n c_i \cdot (\inv{\PrecWorse}_i - \inv{\PrecBetter}_i)} \\
&= \sigmoid \SquareBracket{k(n) \DotProd{c(P)}{\inv{\PrecWorse} - \inv{\PrecBetter}}}
\end{align*}
It is easy to see that $p(\Better{\PrecBetter}{\PrecWorse} | P) > 0.5$ if and only if $C(\PrecWorse | P) > C(\PrecBetter | P)$,
and that the value of $p$ can be increased by increasing $C(\PrecWorse | P)$ or decreasing $C(\PrecBetter | P)$.

We consider the difference $C(\PrecWorse | P) - C(\PrecBetter | P)$
to be the "logit"\todo{MS: Logit je taky tezky pojem, urcite ne automaticky znamy u ATP people.} score of the pair of precedences $\PrecBetter, \PrecWorse$,
and denote it by $C(\Better{\PrecBetter}{\PrecWorse}{P})$.

We define the loss of $c$ on the training sample $(\PrecBetter, \PrecWorse, P)$ as binary cross-entropy
with ground truth $\Better{\PrecBetter}{\PrecWorse}{P}$:

\begin{align*}
L(\PrecBetter, \PrecWorse, P)
&= - \log p(\Better{\PrecBetter}{\PrecWorse}{P}) \\
&= - \log \sigmoid \SquareBracket{k(n) \DotProd{c(P)}{\inv{\PrecWorse} - \inv{\PrecBetter}}} \\
&= - \log \sigmoid \SquareBracket{k(n) \sum_{i=1}^n c_i \cdot (\inv{\PrecWorse}_i - \inv{\PrecBetter}_i)}
\end{align*}

The loss is differentiable with respect to the symbol costs:
\begin{align*}
\frac{\partial L}{\partial c_i}
&= -\sigmoid(-C(\Better{\PrecBetter}{\PrecWorse}{P})) \cdot k(n) \cdot (\inv{\PrecWorse}_i - \inv{\PrecBetter}_i) \\
&= (p(\Better{\PrecBetter}{\PrecWorse}{P}) - 1) \cdot k(n) \cdot (\inv{\PrecWorse}_i - \inv{\PrecBetter}_i)
\end{align*}

### 2021-02-22 evaluation.tex
Reference the GCN description from Evaluation
Mention problem P when comparing precedences

In addition to the signature and the structure of the problem,
some metadata is extracted from the input problem to allow training a more efficient recommender.
First, each clause is annotated with its role in the problem,
which can be either axiom, assumption, or negated conjecture.
% This is simplified. There are actually 7 clause roles defined internally in Vampire,
% and it seems that only 4 of them are ever used in FOL problems:
% AXIOM, ASSUMPTION, CONJECTURE, NEGATED_CONJECTURE
% CONJECTURE is used in FOF problems and NEGATED_CONJECTURE is used in CNF problems,
% so the recommender can actually differentiate between these two forms.
Second, each symbol is annotated with two bits of data:
whether it was introduced into the problem during preprocessing,
and whether the symbol appears in a conjecture clause.
% Note that the second bit is redundant, since the clauses are annotated with their roles.

@@

The two executions are compared in terms of performance:
predicate precedence $\PrecBetter$ is recognized as better than predicate precedence $\PrecWorse$,
denoted as $\Better{\PrecBetter}{\PrecWorse}$,
if the proof search finishes successfully with $\PrecBetter$ ($\SolverRun{P}{\Prec} = \top$)
and if the number of iterations of the saturation loop with $\PrecBetter$ is smaller than with $\PrecWorse$.
\todo{FB: Use notation for saturation loop iterations if the notation was established earlier. 
If failure is denoted by $\infty$, we can simply use $<$.}
If one of the two precedences is recognized as better,
the example $(P, \PrecBetter, \PrecWorse)$ is produced,
where $\PrecBetter$ is the better precedence,
and $\PrecWorse$ is the other precedence.
Otherwise, for example, if Vampire fails on both precedences, we go back to sampling another problem.

>>>>

In addition to the signature and the structure of the problem,
some metadata is extracted from the input problem to allow training a more efficient recommender.
First, each clause is annotated with its role in the problem,
which can be either axiom, assumption, or negated conjecture.
% This is simplified. There are actually 7 clause roles defined internally in Vampire,
% and it seems that only 4 of them are ever used in FOL problems:
% AXIOM, ASSUMPTION, CONJECTURE, NEGATED_CONJECTURE
% CONJECTURE is used in FOF problems and NEGATED_CONJECTURE is used in CNF problems,
% so the recommender can actually differentiate between these two forms.
Second, each symbol is annotated with two bits of data:
whether it was introduced into the problem during preprocessing,
and whether the symbol appears in a conjecture clause.
% Note that the second bit is redundant, since the clauses are annotated with their roles.
This metadata is used to construct the initial embeddings of the respective nodes
in the graph representation of the problem (see \cref{sec:gcn}).
\todo{FB: Consider simplifying the paragraph by removing overlap with \cref{sec:gcn}.}

@@

The two executions are compared in terms of performance:
predicate precedence $\PrecBetter$ is recognized as better than predicate precedence $\PrecWorse$,
denoted as $\Better{\PrecBetter}{\PrecWorse}{P}$,
if the proof search finishes successfully with $\PrecBetter$ ($\SolverRun{P}{\Prec} = \top$)
and if the number of iterations of the saturation loop with $\PrecBetter$ is smaller than with $\PrecWorse$.
\todo{FB: Use notation for saturation loop iterations if the notation was established earlier. 
If failure is denoted by $\infty$, we can simply use $<$.}
If one of the two precedences is recognized as better,
the example $(P, \PrecBetter, \PrecWorse)$ is produced,
where $\PrecBetter$ is the better precedence,
and $\PrecWorse$ is the other precedence.
Otherwise, for example, if Vampire fails on both precedences, we go back to sampling another problem.
